---
title: "Test - extract environmental data for coordinates - package xyzt & friends"
subtitle: "OHW 2022"
date: "`r Sys.Date()`"
output: html_notebook
---

This is a test run using package [xyzt](https://github.com/BigelowLab/xyzt) (and derivatives for specific data sources) to extract e  environmental data for coordinates of interest (e.g. sampling stations).  
I'm using my MSFD dataset for 2017, containing zoobenthic stations in the Bulgarian Black Sea sampled by IO-BAS for the national MSFD monitoring. There are separate coordinates for each replicate sample, sampling dates, and station depths in meters.  


## Setup  
Load necessary packages.  
```{r load_packages}
library(tidyverse)
library(here)
library(sf)
library(ghrsst)
library(xyzt)
library(stars)
```

Define the working directory using package here (= project directory, ohw22-proj-xyzt), by giving it the path to the current file, relative to the project root.  
```{r set_wd}
here::i_am("notebook/test_covars_extraction_with_xyzt.Rmd")
```


## Import data  
Read in my example dataset (in the data subdirectory). It's a csv file.   
```{r import_data}
(msfd <- read_csv(here("data", "msfd_2017_meta.csv")) %>% 
     ## remove units from depth column name 
     rename(depth = depth_m)
)
```

Transform to spatial points - ignore the dates and depths for now.  
```{r make_spatial_points}
(msfd.sp <- msfd %>% 
    as_POINT(dims = "xy")
)
```


## Extract data for points  
Get the environmental data - in this case, SST from [GHRSST (Group for High Resolution Sea Surface Temperature)](https://www.ghrsst.org/). Several other variables are also available, but the time to get them all is fairly long, and I want to go through all dates and the corresponding points in the dataset in one go.    


```{r extract_sst_points}
## get the unique dates
dates <- msfd.sp %>% pull(date) %>% unique()

(msfd.sp.sst <- map(dates, function(x) {
    ## subset points to only those sampled on the current date
    msfd.sub <- msfd.sp %>% 
        filter(date == x)
    
    ## generate a MUR url for the current date
    url <- mur_url(x)
    
    ## open the resource
    X <- ncdf4::nc_open(url)

    ## extract the data - only SST 
    covars <- ghrsst::extract(msfd.sub, X, varname = "analysed_sst")
    
    ## bind to the input
    msfd.sub.sst <- dplyr::bind_cols(msfd.sub, covars)
    
    ## close the resource
    ncdf4::nc_close(X)
    
    return(msfd.sub.sst)
    
}) %>%
        
    ## put the dataset back together in a tibble
    bind_rows()
)
```


Ok, seems fine.. Note how the SSTs look duplicated - because the replicate samples are so close to one another (in fact I wouldn't really record 2 different sets of coordinates for replicates at the same station, but oh well, IO-BAS do this with their samples).  


## Extract data for bounding box (from points)  
Use the points' spatial extent to construct a bounding box, then get the environmental data.  
This time, I'll just use one date, and get all available parameters.  

```{r make_bounding_box}
msfd.bbox <- msfd %>% 
    as_BBOX()
```


Get the environmental data.  
```{r extract_env_data_bounding_box}
## generate a MUR url for the desired date - here, the first one (only one remaining for the bounding box)
url <- mur_url(msfd.bbox %>% pull(date))
    
## open the resource
X <- ncdf4::nc_open(url)

## extract data
(env.data <- ghrsst::extract(msfd.bbox, X, varnames = ghrsst::mur_vars(X)))

## close the connection
ncdf4::nc_close(X)
```


Plot the extracted SST and the points.  
```{r plot_sst_bounding_box}
plot(env.data, attr = 'analysed_sst', axes = TRUE, reset = FALSE)
plot(sf::st_geometry(msfd.sp), add = TRUE, col = "orange", pch = 19, cex = 0.9)
```

