---
title: "Test - extract environmental data for coordinates from Copernicus marine - package copernicus (xyzt)"
subtitle: "OHW 2022"
date: "`r Sys.Date()`"
output: html_notebook
---

This is a test run using package [copernicus](https://github.com/oceanhackweek/ohw22-proj-xyzt/tree/main/package/copernicus) developed during OceanHackWeek2022 to extract environmental data for coordinates of interest (e.g. sampling stations) from [Copernicus marine service](https://marine.copernicus.eu).  
The test script downloads the very first dataset in the catalogue, [GLOBAL_ANALYSIS_FORECAST_PHY_001_024 (Global Ocean 1/12° Physics Analysis and Forecast updated Daily)](https://resources.marine.copernicus.eu/product-detail/GLOBAL_ANALYSIS_FORECAST_PHY_001_024/INFORMATION).  
I'm also using my MSFD dataset for 2017, containing zoobenthic stations in the Bulgarian Black Sea sampled by IO-BAS for the national MSFD monitoring. There are separate coordinates for each replicate sample, sampling dates, and station depths in meters.  


## Setup  
Package copernicus can be installed from Github. See readme for installation instructions and other requirements.  
A (free) Copernicus account is required to download data; the readme explains how to store the credentials in the R environment.   

Load necessary packages.  
```{r load_packages}
library(tidyverse)
library(here)
library(xyzt)
library(copernicus)
```

Define the working directory using package here (= project directory, ohw22-proj-xyzt), by giving it the path to the current file, relative to the project root.  
```{r set_wd}
here::i_am("notebook/test_covars_extraction_with_copernicus.Rmd")
```


## Import data  
Read in my example dataset (in the data subdirectory). It's a csv file.   
```{r import_data}
(msfd <- read_csv(here("data", "msfd_2017_meta.csv")) %>% 
     ## remove units from depth column name 
     rename(depth = depth_m)
)
```

Transform to spatial points - ignore the dates and depths for now.  
```{r make_spatial_points}
(msfd.sp <- msfd %>% 
    as_POINT(dims = "xy")
)
```


## Download Copernicus data  
The package contains a script for constructing the server request for a given dataset and bounding box of interest. The coordinates for the bounding box are in the order **lon_min, lon_max, lat_min, lat_max**. The dates and depth of interest can also be specified.    
The script downloads the whole bounding box, from which the environmental parameters values at whatever coordinates can be extracted.   

Define the bounding box of interest (W Black Sea - somewhat arbitrarily defined here) and download data.  
```{r download_data_copernicus}
## insert the bounding box coordinates in the script (lon_min, lon_max, lat_min, lat_max)  
tst.script <- populate_script(bb = c(27.44547, 30.67082, 41.96027, 44.03979), 
                              out_name = here("data", "copernicus_2022-08-19.nc")) ## put downloaded file in data folder 

download_copernicus(tst.script)
```


Read in the downloaded file.  
```{r import_downloaded_data_copernicus}
(bs <- read_copernicus(here("data", "copernicus_2022-08-19.nc")))
```

Extract the values of the environmental variables at the sampling stations.  
```{r extract_env_data_stations}
(vars <- stars::st_extract(bs, at = msfd.sp))
```

Plot one of the parameters (bottom temperature) - for the whole downloaded area and only for the bounding box defined by the sampling stations.    
```{r plot_downloaded_data}
## define the bounding box of the sampling stations
bbox <- msfd.sp %>% 
    xyzt::as_BBOX()

## subset the downloaded Copernicus dataset to only this bounding box
(sub <- bs[bbox])

## plot the bottom temperature
plot(bs['bottomT'], axes = TRUE, reset = FALSE)
plot(sub['bottomT'], add = TRUE, col = rainbow(6), main = NULL)
```

