---
title: "copernicus"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stars)
library(copernicus)
```

# copernicus

Provides download and access to [Copernicus](https://marine.copernicus.eu/) marine datasets.

### Requirements

#### Python
  
  + [motuclient](https://help.marine.copernicus.eu/en/articles/4796533-what-are-the-motu-client-motuclient-and-python-requirements#h_253e379790)

#### R

  + [R v4+](https://www.r-project.org/)
  + [rlang](https://CRAN.R-project.org/package=rlang)
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  + [ncdf4](https://CRAN.R-project.org/package=ncdf4)
  + [sf](https://CRAN.R-project.org/package=sf)
  + [stars](https://CRAN.R-project.org/package=stars)
  + [readr](https://CRAN.R-project.org/package=readr)
  
#### Copernicus Credentials

You have to get your own credentials from [Copernicus](https://resources.marine.copernicus.eu/registration-form) - a user name and password.  These can be stored in your `~/.Rprofile` file which is just a hidden text file that is loaded and run when you instance of R starts.

```
options(gargle_oauth_email = "btupper@bigelow.org",
        copernicus = list(username = "my_user_name", 
                          password = "my_password"))
```

  
### Installation

```
remotes::install_github("oceanhackweek/ohw22-proj-xyzt",
                        subdir = "package/copernicus")
```

## Downloading

As minimum you should be able to down the default (for Northwest Atlantic). 

```{r}
library(stars)
library(copernicus)

x <- fetch_copernicus() |>
  Reduce(c)
x
```


### Look Up Table (lut)

The variable names are a bit cryptic (at least to me) so we provide a look-up table that associates the variable name with other information available from the source.

```{r lut, message = FALSE}
lut <- copernicus::read_lut()
print(lut, n = nrow(lut))
```


