---
title: "copernicus"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# copernicus

Provides download and access to [Copernicus](https://marine.copernicus.eu/) marine datasets.

### Requirements

#### Python
  
  + [motuclient](https://help.marine.copernicus.eu/en/articles/4796533-what-are-the-motu-client-motuclient-and-python-requirements#h_253e379790)

#### R

  + [R v4+](https://www.r-project.org/)
  + [rlang](https://CRAN.R-project.org/package=rlang)
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  + [ncdf4](https://CRAN.R-project.org/package=ncdf4)
  + [sf](https://CRAN.R-project.org/package=sf)
  + [stars](https://CRAN.R-project.org/package=stars)
  + [readr](https://CRAN.R-project.org/package=readr)
  
#### Copernicus Credentials

You have to get your own credentials from [Copernicus](https://resources.marine.copernicus.eu/registration-form) - a user name and password.  These can be stored in your `~/.Rprofile` file which is just a hidden text file that is loaded and run when you instance of R starts.

```
options(copernicus = list(username = "my_user_name", 
                          password = "my_password",
                          python_app = "python3"))
```

If you have a non-boring install of python then be sure to set python_app to your own need. 

  
### Installation

```
remotes::install_github("oceanhackweek/ohw22-proj-xyzt",
                        subdir = "package/copernicus")
```

## Downloading

As minimum you should be able to down the default (for Northwest Atlantic) region. 

```{r fetch}
library(stars)
library(copernicus)
library(xyzt)

x <- fetch_copernicus(out_path = "~/copernicus-test.nc") 
x
```

You can read that file later using `read_copernicus(filename)` like this.

```{r read}
x <- read_copernicus("~/copernicus-test.nc")
```

### Extracting point data

Since have a stars object, you can use `stars::st_extract(x, at = your_points)`.
Let's borrow the Gulf of Maine points from the [xyzt package](https://github.com/BigelowLab/xyzt)

```{r extract points}
points <- xyzt::read_gom() |>
  xyzt::as_POINT()

vars <- stars::st_extract(x, at = points)
vars
```

### Extracting data within a bounding box

```{r extract bbox}
bbox <- points |>
  xyzt::as_BBOX()

sub <- x[bbox]
sub
```

```{r plot}
n <- 11
plot(x['bottomT'], axes = TRUE, reset = FALSE)
plot(sub['bottomT'], add = TRUE, 
     col = rainbow(n-1),
     main = NULL)
```

### Look Up Table (lut)

The variable names are a bit cryptic (at least to me) so we provide a look-up table that associates the variable name with other information available from the source.

```{r lut, message = FALSE}
lut <- copernicus::read_lut()
print(lut, n = nrow(lut))
```
